### 2020-10-16 papers/ijcar2020/response-to-rebuttal.txt
Move the svn papers to git.  I tried to move the histories also but failed with that.  In the end, I just did a unix cp and git add .

>>  ----------------------- REVIEW 1 ---------------------

>>  ... My understanding is that they have a Coq tactic that takes
>>  a proof goal, exports to to an external mini-prover, which generates
>>  an FPC certificate. Then the FPC certificate is imported back to Coq
>>  via a deep embedding, and Coq uses a reflection theorem to turn it
>>  into an ordinary proof. But I am only guessing: this is not made
>>  clear.

A more accurate outline is: an external proof certificate is checked
via lambda Prolog (ELPI running within Coq).  If that check is
successful, lambda Prolog will build a proof structure that the Coq
kernel should be able to check (only the Coq kernel needs to be
trusted).  Part of the novelty here is the use of FPCs and logic
programming, which makes defining external proof structures, checking
such proofs, and generating Coq proofs all part of the same, modular
process.  Another novelty here is that Coq proofs are built entirely
within the ELPI plugin without using, say, Ltac or reflection.

>>  The example, a proof of "(A -> B) -> A -> B", is singularly
>>  unimpressive. Can the tool do something more complex? How long does it
>>  take?

Since this is an ongoing project, there are now more examples in the
GitHub repository.  Many other kinds of FPCs are described in 
papers [5,9,11].

>>  I don't understand what is in the repository. The READMEs don't
>>  provide any guidance.

The repository has been reorganized to make it much more clear.

>>  ----------------------- REVIEW 2 ---------------------

>>  - The presented example (example 1) that uses the system is
>>    trivial. It leaves me wondering how easy it is to apply the system
>>    to realistic examples.

There are more examples in the GitHub repository.

>>  * I am uncertain of the accuracy of the second sentence in the related
>>    work section. It states that this version of FPC-Coq only works for
>>    proving theorems in first order intuitionistic logic. Then continues
>>    to state that FPC-Coq can be used to prove the double negation of a
>>    formula for which there is a classical proof. Double negating a
>>    formula and finding an intuitionistic proof is only true for
>>    propositional logic and not first-order logic. First order negative
>>    translations are more involved (e.g. also double negating the body
>>    of each forall quantifier).

We meant to say "double negation translation" and not "the double
negation."  As this reviewer points out, many more negations need to
be inserted into first-order formulas for this scheme to work.  The
use of such double-negation translations in this focused setting has
been studied more carefully in reference [11] and in the technical
report https://hal.inria.fr/hal-01354298 (by Chihani, Ilik, Miller).

>>  - Section 2.1, could you please provide a short explanation why there
>>    are 101 introduction rules (and not 100)?

Since there are 100 formulas on the left and one on the right, there
are possibly 101 places to use an introduction rule.

>>  - Section 2.2: it may be helpful to explicitly state that decide and
>>    store are examples of experts and not necessarily part of every
>>    system.

We can be more clear about this: Decide and store are part of the
focusing mechanism and, hence, part of every FPC.  They are not,
however, part of the usual presentation of the sequent calculus.