### 2020-10-16 papers/coq-workshop-2020/main.tex

\documentclass[a4paper]{easychair}
\usepackage{xspace}
\usepackage{fullpage}

\newcommand{\lP}{$\lambda$Prolog\xspace}
\newcommand{\LKF}{\hbox{\sl LKF}\xspace}
\newcommand{\LJF}{\hbox{\sl LJF}\xspace}
\newcommand{\fpccoq}{FPC-Coq\xspace}

\title{\fpccoq: Using ELPI to elaborate external proof evidence into Coq proofs}
\author{Roberto Blanco\inst{1}\and Matteo Manighetti\inst{2}\and Dale Miller\inst{2}}
\institute{Inria Paris\and Inria Saclay}
% RB: I wonder whether/when my change of affiliation will be relevant.
\authorrunning{Blanco, Miller, and Manighetti}
\titlerunning{\fpccoq: Using ELPI in Coq}

\begin{document}
\maketitle
\begin{abstract}\noindent
Logic programming implementations of the \emph{Foundational Proof
  Certificate} (FPC) framework are capable of checking a wide range of
proof evidence.  Proof checkers based on logic programming can make
use of both unification and backtracking search to allow varying
degrees of proof reconstruction to take place during proof checking.
Such proof checkers are also able to elaborate proofs lacking full
details into proofs containing much more detail.  We outline here a
proof-of-concept system in which the Coq-Elpi plugin, which embeds an
implementation of \lP into Coq, is used to check proof certificates
supplied by external (to Coq) provers and to elaborate them into the
fully detailed proof terms that can be checked by the Coq kernel.
\end{abstract}

The trusted base of Coq is its kernel, which is a type-checking
program that certifies that a dependently typed $\lambda$-term has a
given type.  If type checking succeeds, we accept that the formula
corresponding to that type is, in fact, a theorem of intuitionistic
logic.  The rest of the Coq system, especially its tactic language, is
designed to help a human user build proofs-cum-$\lambda$-terms that
can be checked by the kernel.  Of course, there are many theorem
provers for intuitionistic logic for which a completed proof is not
the kind of detailed $\lambda$-term required by the Coq kernel but is,
instead, a trace of some key aspects of a proof: some proof details
might not be captured in such a trace.  For example:
\begin{enumerate}

\item Substitution instances of quantifiers might not be recorded in a
  proof since such instances can, in principle, be reconstructed using
  unification.

\item Detailed typing information might not need to be stored within a
  proof certificate since types can often be reconstructed during
  proof checking.

\item Some simplifications steps might be applied within a proof
  without recording which rewrites were used.  A simple 
  non-deterministic proof-search engine might be expected to
  reconstruct an equivalent simplification.
\end{enumerate}
%RB: Not intuitionistic, but SAT/SMT certificates also are like this.

The Foundational Proof Certificate (FPC) framework \cite{chihani17jar}
has been designed for formally defining a rich collection of proof
certificates that do not necessarily contain full details.  Proof
checkers for FPC certificates can be written using simple logic
programs: the standard implementation mechanisms of unification and
backtracking search can then be used to reconstruct missing proof
details \cite{miller17fac}.  Such proof checking kernels can also be
used to elaborate such certificates into fully detailed proofs
\cite{blanco17phd,blanco17cade}.

We built the \fpccoq system to demonstrate the feasibility of using
logic programming to check proof certificates produced by external
provers and to elaborate such certificates into proofs with sufficient
details that they can be given directly to Coq's kernel to
check.  Note that one does not need to trust the external prover nor
the \fpccoq tool-chain: one only needs to trust the Coq kernel.

Because of the need to represent and compute with quantificational
formulas and dependently-typed $\lambda$-terms, we chose \lP
\cite{miller12proghol} as our implementation language since it has a
direct and elegant treatment of binding in data structures.  The ELPI
implementation \cite{dunchev15lpar} of \lP has been embedded recently
into the Coq-Elpi plugin \cite{coq-elpi20web}, and we have used this
plugin to implement a system that checks proof certificates in the FPC
framework and elaborates them into proof terms that can be submitted
to the Coq kernel.

The \fpccoq system can be seen as part of the larger effort to use \lP
within a type theory setting \cite{coen19mscs}, in general, and within
Coq \cite{tassi19itp}, in particular. In doing so, the first step is to
define how to use Coq terms as proof evidence for a formula, i.e.,
to program a type checker in the FPC framework. Initially, the
presentation of connectives as inductive types in Coq is translated to
the usual connectives of intuitionistic first-order logic, the only ones
known to the FPC checker.

The current version of \fpccoq only works for proving theorems in
first-order intuitionistic logic.  Given that the \lP proof checker,
based on \LJF (a focused version of intuitionistic sequent calculus),
can also serve as a proof checker for \LKF (its classical
counterpart), \fpccoq can be used to prove double-negation
translations of a formula for which there is a proof certificate in
classical logic \cite{chihani17jar}.  Extending the FPC framework to
involve inductive and coinductive reasoning has also been considered
\cite{heath15pxtp,heath19jar}, and a future version of \fpccoq could
well include proof certificates that are output from inductive theorem
provers as well as model checkers.

Given its origins in the theory of focused sequent calculus, our first
FPC proof checker only implements a minimalistic sequent calculus in a
purely declarative fashion.  We have also developed a second approach
to building a proof checker based on dependent types in the style of
\cite{lengrand10lmcs}.  This second approach allows for a much more
concise implementation, where there is no translation needed from Coq
formulas to intuitionistic formulas since the kernel directly operates
with dependent types. The usual definitions of proof certificates
remain directly usable also in this context.

In particular, when the user loads the \lP code for an FPC definition
named, say \texttt{fpc}, into the system, that FPC definition is made
available as a Coq-Elpi tactic that takes a proof certificate, say
\texttt{cert}, as an argument.  Inside Coq's proof mode, that tactic
can be invoked by \texttt{elpi fpc cert}.

The code for both of these implementations is available at
\url{https://github.com/proofcert/fpc-elpi}, along with some simple
examples to illustrate the steps taken to elaborate and check proof
certificates.  We have not yet explored the cost of performing proof
checking in this setting: it is likely to vary greatly among different
certificate formats.

Although \fpccoq currently targets the construction of proofs for the
Coq kernel, it should be straightforward to use the same setup
described here to build Dedukti proofs as well \cite{assaf16types}.
Our development could also provide the facilities for cleanly and
easily building Coq tactics that are written in \lP and are controlled
by (partial) proof certificates.

%\bibliography{../../references/master}
\bibliographystyle{plain}
\begin{thebibliography}{10}

\bibitem{assaf16types}
Ali Assaf, Guillaume Burel, Raphal Cauderlier, David Delahaye, Gilles Dowek,
  Catherine Dubois, Fr{\'e}d{\'e}ric Gilbert, Pierre Halmagrand, Olivier
  Hermant, and Ronan Saillard.
\newblock Expressing theories in the $\lambda$${\Pi}$-calculus modulo theory
  and in the {Dedukti} system.
\newblock In {\em TYPES: Types for Proofs and Programs}, Novi Sad, Serbia,
  2016.

\bibitem{blanco17phd}
Roberto Blanco.
\newblock {\em Applications for Foundational Proof Certificates in theorem
  proving}.
\newblock PhD thesis, Universit{\'e} Paris-Saclay, December 2017.

\bibitem{blanco17cade}
Roberto Blanco, Zakaria Chihani, and Dale Miller.
\newblock Translating between implicit and explicit versions of proof.
\newblock In Leonardo de~Moura, editor, {\em Automated Deduction - {CADE} 26 -
  26th International Conference on Automated Deduction, Gothenburg, Sweden,
  August 6-11, 2017, Proceedings}, volume 10395 of {\em Lecture Notes in
  Computer Science}, pages 255--273. Springer, 2017.

\bibitem{chihani17jar}
Zakaria Chihani, Dale Miller, and Fabien Renaud.
\newblock A semantic framework for proof evidence.
\newblock {\em J. of Automated Reasoning}, 59(3):287--330, 2017.

\bibitem{dunchev15lpar}
Cvetan Dunchev, Ferruccio Guidi, Claudio~Sacerdoti Coen, and Enrico Tassi.
\newblock {ELPI:} fast, embeddable, $\lambda${Prolog} interpreter.
\newblock In M. Davis, A. Fehnker, A. McIver, and A. Voronkov, editors, 
{\em Logic for Programming, Artificial Intelligence, and
  Reasoning, LPAR-20}, LNCS 9450, pages 460--468. Springer, 2015.

\bibitem{coen19mscs}
Ferruccio Guidi, Claudio~Sacerdoti Coen, and Enrico Tassi.
\newblock Implementing type theory in higher order constraint logic
  programming.
\newblock {\em Mathematical Structures in Computer Science}, 29(8):1125--1150,
  2019.

\bibitem{heath15pxtp}
Quentin Heath and Dale Miller.
\newblock A framework for proof certificates in finite state exploration.
\newblock In Cezary Kaliszyk and Andrei Paskevich, editors, {\em Proceedings of
  the Fourth Workshop on Proof eXchange for Theorem Proving}, number 186 in
  Electronic Proceedings in Theoretical Computer Science, pages 11--26. Open
  Publishing Association, August 2015.

\bibitem{heath19jar}
Quentin Heath and Dale Miller.
\newblock A proof theory for model checking.
\newblock {\em J. of Automated Reasoning}, 63(4):857--885, 2019.

\bibitem{lengrand10lmcs}
{St\'ephane} Lengrand, Roy Dyckhoff, and James McKinna.
\newblock A focused sequent calculus framework for proof search in pure type
  systems.
\newblock {\em Logical Methods in Computer Science}, 7(1), 2011.

\bibitem{miller17fac}
Dale Miller.
\newblock Proof checking and logic programming.
\newblock {\em Formal Aspects of Computing}, 29(3):383--399, 2017.

\bibitem{miller12proghol}
Dale Miller and Gopalan Nadathur.
\newblock {\em Programming with Higher-Order Logic}.
\newblock Cambridge Univ. Press, 2012.

\bibitem{tassi19itp}
Enrico Tassi.
\newblock {Deriving Proved Equality Tests in Coq-Elpi: Stronger Induction
  Principles for Containers in Coq}.
\newblock In John Harrison, John O'Leary, and Andrew Tolmach, editors, {\em
  10th International Conference on Interactive Theorem Proving (ITP 2019)},
  volume 141 of {\em Leibniz International Proceedings in Informatics
  (LIPIcs)}, pages 29:1--29:18, Dagstuhl, Germany, 2019. Schloss
  Dagstuhl--Leibniz-Zentrum fuer Informatik.

\bibitem{coq-elpi20web}
Enrico Tassi.
\newblock Coq plugin embedding {ELPI}.
\newblock \url{https://github.com/LPCIC/coq-elpi}, 2020.

\end{thebibliography}

\end{document}

%%  LocalWords:  ELPI Coq's minimalistic Dedukti Elpi fpc elpi