### 2020-10-16 papers/ijcar2020/outtakes.tex

which is capable to bridging the
gap between such high-level and partial proof evidence supplied by
some external prover and the specific needs of the Coq kernel.
%
%
The Foundational Proof Certificate (FPC) framework can use
the unification and backtracking mechanisms of \lP to perform proof
reconstruction in order to provide a fully detailed natural deduction
proof that can be checked by the Coq kernel.
%
As a result, we are able to have Coq check proof evidence supplied by
various external provers in such a way that we do not need to trust
either the external prover or the Coq-ELPI system.