### 2020-10-16 papers/coLP/long-related.tex

\paragraph{Haskell}

QuickCheck can partially deal with infinite objects, provided that the properties themselves are computable. Lazy evaluation is no magic bullets: stream equality loops, but we can use the \emph{take} lemma to observe finite portions.

\paragraph{coLP}

Gupta and his students have been advocating (very, very incrementally)
for extending LP first only coinduction, then allowing both least and
greatest semantics~\cite{Luke07}. The approach is semantically based:
you start with the greatest complete Herbrand model, containing finite
and infinite (both rational and irrational) ground terms. Then they
modify the operational semantics dropping the occur check and relying
on cycle (same atom) and loop detection (same atom modulo
unification). They show soundness of the procedure wrt the greatest
complete Herbrand model. Completeness seems hopeless for
recursion-theoretic reasons as shown in~\cite{AnconaD15}.

\paragraph{Katya}
This has been much improved by Komendantskaya and friends (see~\cite{BasoldKL19}) \dots

Rather than relying on infinite terms via rational unification, they
introduce in the language of the logic a fixed point constructor
(which is then restricted to occur guarded, so as to be normalizing)
and its equational theory. While this may make sense for programming,
it seems problematic for a logical framework. Do we still have
adequacy of the encodings and canonical forms? Do \texttt{fix z.~cons
  0 z} and \texttt{fix z.~cons 0 (cons 0 z)} denote the same infinite
term or now the encodings refers to equivalence classes (perhaps wrt
bi-similarity) ? It would be hard to give a Curry-Howard reading of
this, since fixed points terms do not have a type classifying them and
the later modality does not have a proof term. Still, it could be
straighten up following~\cite{abelVezzosi:aplas14}. However, the
approach seems limited to guarded coinduction and may suffer from the
very problems raised by Ancona et al., namely  that a greatest fixed point semantics plus infinite (although) guarded terms gives unwanted results. For example, I think that  this is
still provable: \texttt{last (fix ones.~cons 1 ones) 2}. Finally, it's hard to implement in our setup, since we have to change the unification algorithm to take into account unfolding of fixed points (it's equi-recursive, in Peirce's terminology).

\paragraph{Terms Generation}

There's been considerable work on generation (mostly random) of
lambda-terms, with an emphasis on uniform
distributions~\cite{BendkowskiGT17}. Tarau generalizes it to arbitrary
first-order signatures~\cite{Tarau18} extending Remy's algorithm for
generating random binary trees \dots

\paragraph{Coaxioms}

In a lengthy list of papers such as ~\cite{AnconaDZ17, Dagnino19},
Ancona at al.\ argue against the use of gfixpoints as the semantics of
Aczel's style inference system. Starting from the desire to improve on
Leroy's coinductive operational semantics given its shortcomings, as
well as those of co-logic programming, they try to carve out undesired
results: take for example the logic program for last element of a list
and an infinite list of ones \texttt{Os = [1|Os]}. Then by a circular
proof you can prove \texttt{last Os 2}. We would argue that (1) we do
not care about infinite objects (2) in a guarded sense, the definition
of \texttt{last} is \emph{not productive} and it is rejected by Coq
(as a \textbf{Fixpoint} (but that proof holds with a
\textbf{Coinductive}) and by Agda's copatterns. They instead use
coaxioms (and then corules) to rule out those derivations. Those are
standard rules, but with a different interpretation: an atom is
provable if it has an infinite (even irrational ) proof, but whose
nodes have all a finite proof from the coaxioms only. Semantically
this corresponds to a fixed point that lies somewhere between least
and greatest. In the \texttt{last} example, the coaxiom would state
\texttt{last [X|XS] X}, which should force the element to be in the
list and rule out the offending derivation.  To me this notion of a
proof does not seem effective (how do I check the coaxiom condition
finitely?) and the coaxioms do not seem to have a declarative reading,
or, less charitably, that reading makes no sense. Still, they may have
more details in later papers. Essentially, it's an answer to a problem
I do not much care for.

\paragraph{Abella's current setup}

I think we all agree that infinite terms are an hassle, better to stay
clear from. Still, Abella's approach makes me worry: first, there is
no formal connection between the guarded version implemented and the
fixed point rules of the meta-logic, in particular compared to the strict correspondence exhibited by Beluga's script language
\emph{Harpoon}.  This can be fixed, I assume. While we cannot have a
``paradox'' such as the \texttt{last} example just because we have no
way to construct such a term, there are others mentioned by Ancona
(graph reachability) that are raised by finite terms. Moreover,
examples in the library such as \texttt{colist.thm} are bizarre,
since we prove properties of objects that we cannot construct. So, we fail to prove that
stream equality entails free equality,  but we cannot exhibit two streams that
refute the property.
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "outline"
%%% End:

%  LocalWords:  Tarau Remy's Coaxioms Ancona gfixpoints Aczel's coLP
%  LocalWords:  coinductive Fixpoint Agda's coaxioms corules coaxiom
%  LocalWords:  Herbrand Katya Komendantskaya QuickCheck