### 2020-10-16 papers/coLP/outline.tex

\RequirePackage[utf8]{inputenc}
\RequirePackage[T1]{fontenc}
\documentclass[a4paper]{easychair}
\usepackage{url,xcolor}
\usepackage{proof}
\usepackage{listings}
\usepackage{letltxmacro}

\newcommand{\instan}[1]{\hbox{\sl grnd}~(#1)}
\newcommand{\Pscr}{{\mathcal P}}

\newcommand{\ok}{\checkmark}
% DM The use of pifont and txfonts broke typesetting for me: equality
% signs never appearred and some parentheses did not appear either.
%\newcommand{\noc}{\ding{56}}
\newcommand{\noc}{\ddag}
\newcommand{\lP}{$\lambda$Prolog\xspace}
\newcommand{\nat }{\hbox{\sl nat}\xspace}
\newcommand{\plus}{\hbox{\sl app}\xspace}
\newcommand{\lst}{\hbox{\sl lst}\xspace}
\newcommand{\sort}{\hbox{\sl sort}}
\newcommand{\rev}{\hbox{\sl rev}}
\newcommand{\ordered}{\hbox{\sl ordered}}

%\newcommand{\atac}{A2Tac\xspace}
\newcommand{\atac}{ACheck\xspace}
%\newcommand{\lra}{\mathrel{\longrightarrow}}
\newcommand{\lra}{\mathrel{\vdash}}
\newcommand{\seq}[2]{#1\lra #2}

\newcommand{\true}{tt}

%%%%%%%%%%%%%%%% LJF
\newcommand{\truen}{t^-}
\newcommand{\truep}{t^+}
\newcommand{\falsen}{f^-}
\newcommand{\falsep}{f^+}
\newcommand{\wedgep}{\wedge^{\!+}}
\newcommand{\wedgen}{\wedge^{\!-}}
\newcommand{\veep}{\vee^{\!+}}
\newcommand{\veen}{\vee^{\!-}}
\newcommand{\with}{\&}

\newcommand{\Nscr}{{\cal N}}
\newcommand{\Rscr}{{\cal R}}                                   % Used for an ambiguous rhs
%\newcommand{\Rscr}{\Delta_1\Downarrow\Delta_2}                % Used for an ambiguous rhs
\newcommand{\jUnf    }[4]{#1\mathbin\Uparrow#2\vdash#3\mathbin\Uparrow #4} % unfocused sequent
\newcommand{\jUnfG   }[2]{\jUnf{\Gamma}{#1}{#2}{{}}}           % unf sequ with \Gamma
\newcommand{\jUnfamb }[3]{#1\mathbin\Uparrow#2\vdash#3 \Rscr}  % unfocused sequent
\newcommand{\jUnfGamb}[1]{\Gamma\mathbin\Uparrow#1\vdash \Rscr}% unf sequ with \Gamma
\newcommand{\jLf     }[3]{#1\Downarrow#2\vdash#3}              % left focused sequent
\newcommand{\jLfG    }[1]{\jLf{\Gamma}{#1}{E}}                 % left foc seq with \Gamma 
\newcommand{\jRf     }[2]{#1\vdash #2\Downarrow}               % right focused sequent
\newcommand{\jRfG    }[1]{\jRf{\Gamma}{#1}}                    % right foc seq with \Gamma
%%%%%%%%%%%%%%%% 
\newcommand{\bxi}[1]{\blue{\Xi_{#1}}}
\newcommand{\bXi}[1]{\blue{\Xi_{#1} :\null}}

\newcommand{\andClerk}[3]{{\wedge_c}(#1,#2,#3)}
\newcommand{\falseClerk}[2]{f_c(#1,#2)}
\newcommand{\orClerk}[2]{{\vee_c}(#1,#2)}
\newcommand{\allClerk}[2]{\forall_c(#1,#2)}
\newcommand{\storeClerk}[3]{\hbox{\sl store}_c(#1,#2,#3)}

\newcommand{\trueExpert }[1]{{\true_e}(#1)}
\newcommand{\eqExpert }[1]{{=_e}(#1)}
\newcommand{\unfoldExpert}[2]{{\hbox{\sl unfold}_e}(#1,#2)}
\newcommand{\andExpert}[3]{{\wedge_e}(#1,#2,#3)}
\newcommand{\andExpertLJF}[6]{{\wedge_e}(#1,#2,#3,#4,#5,#6)}
\newcommand{\orExpert  }[3]{{\vee_e}(#1,#2,#3)}
\newcommand{\someExpert}[3]{\exists_e(#1,#2,#3)}
\newcommand{\initExpert}[2]{\hbox{\sl init}_e(#1,#2)}
\newcommand{\cutExpert}[4]{\hbox{\sl cut}_e(#1,#2,#3,#4)}
\newcommand{\decideExpert}[3]{\hbox{\sl decide}_e(#1,#2,#3)}
\newcommand{\releaseExpert}[2]{\hbox{\sl release}_e(#1,#2)}
%%%%%%%%%%%%%%%%

%
\newcommand{\step}{\longrightarrow}
\newcommand{\sstlc}{\textit{STLC}\xspace}
\newcommand{\vds}{\vdash_\Sigma}

\def\bnfas{\mathrel{::=}}
\def\bnfalt{\mid}

\def\lam{\lambda}
\def\Lam{\Lambda}
\def\arrow{\rightarrow}
\def\oftp{\mathord{:}}
\def\hastype{\mathrel{:}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Editorials
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\long\def\ednote#1{\footnote{[{\it #1\/}]}\message{ednote!}}
%\long\def\ednote#1{\begin{quote}[{\it #1\/}]\end{quote}\message{note!}}
\newenvironment{metanote}{\begin{quote}\message{note!}[\begingroup\it}%
                         {\endgroup]\end{quote}}
\long\def\ignore#1{}

\newcommand{\todo}[1]{\begin{metanote}TODO: #1\end{metanote}}
%
%\setlength{\textwidth}{13.2cm}
%\setlength{\textheight}{19.9cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{types20/listing-macros}  % Use to get better highlighing of code
\lstset{language=lprolog}
\lstset{language=abella}
\title{On the Proof Theory of Property-Based Testing of Coinductive Specifications, or:\\
 PBT to Infinity and beyond}
\authorrunning{Blanco, Miller and Momigliano}
\titlerunning{PBT to Infinity and
   beyond}
\author{Roberto Blanco\inst{1} \and Dale Miller\inst{2} \and Alberto Momigliano\inst{3}}

\institute{
  INRIA Paris, France
  \and
  INRIA Saclay \& LIX, \'Ecole Polytechnique, France\\
\and
DI, Universit\`a degli Studi di Milano, Italy
}
\begin{document}

\maketitle

\section{Introduction}
\label{sec:intro}

\begin{center}
  [For the moment taken from the TYPES abstract, to be revised]
\end{center}

Reasoning about infinite computations via coinduction and corecursion
has an ever increasing relevance in formal methods and, in particular,
in the semantics of programming languages, starting
from~\cite{milner91tcs}; see also~\cite{2007-Leroy-Grall} for a
compelling example --- and, of course, coinduction underlies (the
meta-theory of) process calculi. This was acknowledged by researchers
in proof assistants, who promptly provided support for coinduction and
corecursion from the early 90's on, see~\cite{Paulson97,Gim95types}
for the beginning of the story concerning the most popular frameworks.

It also became apparent that tools that search for
refutations/counter-examples of conjectures prior to attempting a
formal proof are invaluable: this is particularly true in PL theory,
where proofs tend to be shallow but may have hundreds of cases.  One
such approach is \emph{property-based testing} (PBT), which employs
automatic test data generation to try and refute executable
specifications.  Pioneered by \emph{QuickCheck} for functional
programming~\cite{claessen00icfp}, it has now spread to most major
proof assistants~\cite{BlanchetteBN11,QChick}.

% DM I revised the following paragraph.  Did I get the sense right?

In general, PBT does not extend well to coinductive specifications
(an exception being Isabelle's \emph{Nitpick},
% \url{https://isabelle.in.tum.de/dist/doc/nitpick.pdf}
which is, however, a counter-model generator).  A particular
challenge, for example, for \emph{QuickChick} is extending it to work
with Coq's notion of coinductive via \emph{guarded} recursion (which
is generally seen to be an unsatisfactory approach to coinduction). We
are not aware of applications of PBT to other form of coinduction, such as \emph{co-patterns}~\cite{AbelPTS13}.

While PBT originated in the functional programming community, we have
given in a previous paper (\cite{Blanco0M19}) a reconstruction of some
of its features (operational semantics, different flavors of
generation, shrinking) in purely proof-theoretic terms employing the
framework of \emph{Foundational Proof Certificates}~\cite{chihani17jar}: the
latter, in its full generality, defines a range of proof structures
used in various theorem provers such as resolution refutations,
Herbrand disjuncts, tableaux, etc.
%
In the context of PBT, the proof theory setup is much simpler.

\begin{center}
  [etc\dots]
\end{center}

The contributions of this paper are
\begin{itemize}
\item Tighten the link of our previous reconstruction of PBT to the
  proof-theoretic approach of model checking in~\cite{heath19jar} by casting it in the setting of $\mu$MALL
\item Extend PBT beyond the Horn-nabla fragment to Harrop spec, first using only case analysis, then also co-induction
\item Experiment with a multi-level architecture, where FPC only takes
  care of generation and testing  is delegated to a system such as Bedwyr
\item Hopefully provide some examples (CCS, $\pi$ calculus (?), co-evaluation)
\item
  \begin{center}
[others \dots]    
  \end{center}

\end{itemize}

\section{The story so far}
\label{sec:tecap}

\begin{center}
  [Here we recap what we did in PPDP, in the uniform proof style, see fig~\ref{fig:augmented}: the connection with linear logic comes in the next section]
\end{center}

Consider an attempt to find counter-examples to a conjecture of the
form \(\forall x [(\tau(x)\wedge P(x)) \supset Q(x)]\) where $\tau$ is
a typing predicate and $P$ and $Q$ are two other predicates defined
using Horn clause specifications.
%
By negating this conjecture, we attempt to find a (focused) proof of 
\(\exists x [(\tau(x)\land P(x)) \land \neg Q(x)]\).
%
In the focused proof setting, the \emph{positive phase} (where
test cases are generated) is represented by \(\exists x\) and
\((\tau(x)\land P(x))\). 
%
That phase is followed by the \emph{negative phase} (where conjectured
counter-examples are tested) and is represented by \(\neg Q(x)\).
%
%% turns out to be a search for a proof of \(\exists x [(\tau(x)\land
%%   P(x)) \land \neg Q(x)]\) in a focused sequent calculus for
%% (basically) Horn logic, and pre-condition phase, while proving $Q$ can
%% be relegated to deterministic logic programming-like computation (the
%% \emph{negative} phase), interpreting ``$\neg$'' as
%% Negation-as-failure..
%
%

\begin{figure}
\newcommand{\XXi}{{\color{blue}{\Xi}}}
\[
\infer{\XXi\vdash G_1\wedge G_2}
      {\XXi_1\vdash G_1\qquad \XXi_2\vdash G_2\qquad \andExpert{\XXi}{\XXi_1}{\XXi_2}}
\qquad
\infer{\XXi\vdash \true}
      {\trueExpert{\XXi}}
\]
\vskip -6pt
\[
\infer{\XXi\vdash G_1\vee G_2}
      {\XXi'\vdash G_i\qquad \orExpert{\XXi}{\XXi'}{i}}
\qquad
\infer{\XXi\vdash \exists x. G}
      {\XXi'\vdash G[t/x]\qquad \someExpert{\XXi}{\XXi'}{t}}
\]
\vskip -6pt
\[
\infer{\XXi\vdash t = t}
      {\eqExpert{\XXi}}
\qquad
\infer{\XXi\vdash A}
      {\XXi'\vdash G \quad (A~\hbox{\tt :-}~G)\in\instan\Pscr
                     \quad \unfoldExpert{\XXi}{\XXi'}}
\]
\caption{A proof system augmented with proof certificates and
  additional ``expert'' premises.}
\label{fig:augmented}
\end{figure}

\section{Linear logic as the proof theory of model checking}
\label{sec:ll}

A natural choice is linear logic cousin $\mu\mathrm{MALL}$~\cite{Baelde12}, which is associated to  the \emph{Bedwyr} model-checker. %  In fact, the latter
% has already been used for related aims~\cite{HeathM15}.

\subsection{Horn programs as fixed points}
\label{ssec:lpfixed}

\begin{center}
  [Here we introduce fixed points, and re-state the rules in fig~\ref{fig:augmented} with only unfolding rules. We may also want to introduce our first coinductive examples, probably \texttt{coeval} (it's Horn). 
  ]
\end{center}

\subsection{Onwards to linear logic}
\label{ssec:ll}

\begin{center}
  [
  \begin{itemize}
  \item Here we introduce polarization, focusing. I assume we will be
    using only $\mu$F rather than all muMALL; and if so, perhaps we
    can simplify the syntax of the focusing judgment
  \item Not sure we want to put clercks and experts everywhere, if they play a part only during generation and that should stay in the Horn+nabla fragment
  \item Now we can introduce non Horn examples such as simulation and points out that the left unfolding rules do case analysis.
\end{itemize}
  ]
\end{center}

\subsection{Full system}
\label{ssec:fullmumall}

\begin{center}
  [Here we add the rules for (co)induction, init (and cut?). In the end we take the muFlogic from the JAR paper and the ``completeness'' results thereof]
\end{center}

\section{Implementation and case studies}
\label{sec:case}

\begin{center}
  [Here we discuss the ``architecture'', namely using the lightweight FPC for generators and Bedwyr for testing]
\end{center}  
\subsection{Examples}
\label{ssec:exs}

\begin{center}
  [Which esamples? The ones which work]
\end{center}

\subsection{Refinements}
\label{ssec:refs}

\begin{center}
  [Anything else to make this interesting: integration with
  proof-outlines, pre-computation of terms 'a la Tarau, full
  ``post-mortem'' FPC, that is also for the testing phase, to
  elaborate further the results, for examples for explanations etc., wild ideas about generating infinite terms with observation functions as in \emph{Hipster} \cite{EinarsdottirJP18}]
\end{center}

\section{Related work}
\label{sec:rel}

\begin{center}
  [to much details, just for my own sake, only a small portion will make it in the paper]
\end{center}
  While coinductive logic programming,
see~\cite{Luke07} and~\cite{BasoldKL19} for a much more principled and
in depth treatment, may at first seem to fit the bill, the need to
model infinite \emph{behavior} rather than infinite objects, that is (ir)rational terms on the
domain of discourse, has lead us to adopt a much stronger logic (and
associated proof theory) with explicit rules for induction and
coinduction.
\input{long-related}
\section{Conclusion}
\label{sec:conc}

\bibliographystyle{abbrv}
\bibliography{colp,../ppdp19/l}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

%  LocalWords:  Coinductive QuickCheck utf inputenc fontenc Herbrand
%  LocalWords:  corecursion coinductive corecursive QuickChick CBV ir
%  LocalWords:  disjuncts PCFL LTS Harrop PBT Coq's coinductively CCS
%  LocalWords:  intuitionistic FPCs Bedwyr Bedwyr's nabla