### 2020-10-21 papers/types2020/preamble.tex

%This is a template for producing LIPIcs articles. 
%See lipics-manual.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the mandatory bibstyle

%TODO mandatory, please use full name; only 1 author per \author
%macro; first two parameters are mandatory, other parameters can be
%empty. Please provide at least the name of the affiliation and the
%country. The full address is optional

\author{Matteo Manighetti}
       {Inria, Palaiseau, France \and LIX, Ecole Polytechnique, France
         \and \url{http://www.lix.polytechnique.fr/Labo/Matteo.Manighetti/}}
       {matteo.manighetti@inria.fr}
       {}
       {}

\author{Dale Miller}
       {Inria, Palaiseau, France \and LIX, Ecole Polytechnique, France
         \and \url{http://www.lix.polytechnique.fr/Labo/Dale.Miller/} }
       {dale.miller@inria.fr}
       {}%https://orcid.org/0000-0003-0274-4954
       {}

\author{Alberto Momigliano}
       {Dipartimento di Informatica, Universit\`a degli studi di Milano, Italy \and
       \url{http://momigliano.di.unimi.it/}}
       {momigliano@di.unimi.it}
       {}
       {}

\authorrunning{M. Manighetti, D. Miller, and A. Momigliano}
 %TODO mandatory. First: Use abbreviated first/middle names. Second
 %(only in severe cases): Use first author plus 'et al.'

\Copyright{M. Manighetti, D. Miller, and A. Momigliano}
%TODO mandatory, please use full first names. LIPIcs license is
%"CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} 
%TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Proof assistant, logic programming, Coq, \lP}
%TODO mandatory; please add comma-separated list of keywords

\category{} %optional, e.g. invited paper

\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversion{A full version of the paper is available at \url{...}.}

\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

%\acknowledgements{I want to thank \dots}%optional

\nolinenumbers %uncomment to disable line numbering

\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%