### 2020-10-26 papers/types2020/main.tex

%% Important dates:
%%  Paper submission: 19 October 31 October 2020
%%  Author notification: 18 January 2021
%%  Final version:  15 February 2021
%%  Publication (presumably): end of March 2021

%% Style requirements of LIPIcs.  The recommended length of a paper is
%% 12-15 pages, excluding front-page(s) (authors, affiliation, keywords,
%% abstract, ...), bibliography and an appendix of max 5 pages. If you
%% need more pages, please ask the editors.

%% Papers have to be submitted in pdf through EasyChair:
%% https://easychair.org/conferences/?conf=types2020postproceed

\documentclass[a4paper,USenglish,cleveref, autoref, thm-restate]{lipics-v2019}
\input head
\input preamble

\title{Interfacing logic programming and Coq} % Other suggestions?
\titlerunning{Interfacing logic programming and Coq}

\begin{document}
\maketitle

\begin{abstract}
We do stuff
\end{abstract}

\input intro
\input cultures
\input sec3
\input pbt
\input elab

\section{Conclusion}

\bibliography{l}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: