### 2020-10-26 papers/types2020/app.tex
appendix

\section{Appendix}
\label{sec:app}

Stuff we may not have space for in the paper
\begin{itemize}
\item The LJF sequent rules
\item the code for
  \lsti{dep_pbt}

\item \dots
\end{itemize}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

% LocalWords:  LJF