### 2020-10-29 papers/types2020/app.tex
some work on sec 3 and 4

\section{Appendix}
\label{sec:app}

Stuff we may not have space for in the paper

In Fig.~\ref{fig:interp} we give a proof system for Horn logic and its encoding in coq-elpi
\todo{explain the interface, quotations etc}

\begin{figure}
\newcommand{\XXi}{\Pscr}
\newcommand{\bc}[2]{\Pscr\Downarrow #1 \vdash #2}
\[
\infer{\XXi\vdash t = t}
{%\eqExpert{\XXi}
}
\qquad
\infer{\XXi\vdash \true}
{%\trueExpert{\XXi}
}
\qquad
\infer{\XXi\vdash G_1\vee G_2}
{\XXi\vdash G_i%\qquad \orExpert{\XXi}{\XXi'}{i}
}
\qquad
\infer{\XXi\vdash \exists x. G}
{\XXi\vdash G[t/x]%\qquad \someExpert{\XXi}{\XXi'}{t}
}
\qquad 
\infer{\XXi\vdash G_1\wedge G_2}
{\XXi\vdash G_1\quad \XXi\vdash G_2%\quad \andExpert{\XXi}{\XXi_1}{\XXi_2}
}
\]
\vskip -18pt
\[
\infer{\XXi\vdash A}
      {\bc D A \quad D \in \Pscr}\qquad
  \infer{\bc A A}{}{}\qquad
  \infer{\bc{\forall x.\, D} A}{\bc{D[t/x]} A}{}\qquad
  \infer{\bc{G\supset D} A}{\bc D A\quad \XXi\vdash G}{}
\]
In the judgment $\bc{D}{A}$, the variable $A$ ranges over only atomic
formulas.
\lstinputlisting[linerange={interp-end}]{code/dep-kernel.mod}
\caption{A proof system for Horn logic and its encoding in coq-elpi.}
\label{fig:interp}
%\lstinputlisting[linerange={sigs-end}]{code/dep-kernel.sig}
%\lstinputlisting[linerange={check-end}]{code/dep-kernel.mod}
%\caption{A dependent kernel for Coq terms.}
%\label{fig:kernel}
\end{figure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

% LocalWords:  LJF