### 2020-10-29 papers/types2020/app.tex
added some related and future work, suggestions for title, revisions to sec 4

\section{Appendix}
\label{sec:app}

In this appendix we list some definitions and pieces of code that may be helpful

\subsection{Interp}

In Fig.~\ref{fig:interp} we the encoding of the vanilla meta-interpreter used in
the testing phase of the \lsti{dep_pbt} tactic. Differently from Fig.~\ref{fig:augmented} we appeal, via quotations, to Coq's defined connectives --- note the use of a green cut to avoid overlapping with the \lsti{Inductive} atomic case and the absence of certificates.

\begin{figure}
% \newcommand{\XXi}{\Pscr}
% \newcommand{\bc}[2]{\Pscr\Downarrow #1 \vdash #2}
% \[
% \infer{\XXi\vdash t = t}
% {%\eqExpert{\XXi}
% }
% \qquad
% \infer{\XXi\vdash \true}
% {%\trueExpert{\XXi}
% }
% \qquad
% \infer{\XXi\vdash G_1\vee G_2}
% {\XXi\vdash G_i%\qquad \orExpert{\XXi}{\XXi'}{i}
% }
% \qquad
% \infer{\XXi\vdash \exists x. G}
% {\XXi\vdash G[t/x]%\qquad \someExpert{\XXi}{\XXi'}{t}
% }
% \qquad 
% \infer{\XXi\vdash G_1\wedge G_2}
% {\XXi\vdash G_1\quad \XXi\vdash G_2%\quad \andExpert{\XXi}{\XXi_1}{\XXi_2}
% }
% \]
% \vskip -18pt
% \[
% \infer{\XXi\vdash A}
%       {\bc D A \quad D \in \Pscr}\qquad
%   \infer{\bc A A}{}{}\qquad
%   \infer{\bc{\forall x.\, D} A}{\bc{D[t/x]} A}{}\qquad
%   \infer{\bc{G\supset D} A}{\bc D A\quad \XXi\vdash G}{}
% \]
% In the judgment $\bc{D}{A}$, the variable $A$ ranges over only atomic
% formulas.
 \lstinputlisting[linerange={sigs-end}]{code/dep-kernel.sig}
\lstinputlisting[linerange={interp-end}]{code/dep-kernel.mod}
\caption{Encoding in coq-elpi of the vanilla meta-interpreter}
\label{fig:interp}
%\lstinputlisting[linerange={sigs-end}]{code/dep-kernel.sig}
%\lstinputlisting[linerange={check-end}]{code/dep-kernel.mod}
%\caption{A dependent kernel for Coq terms.}
%\label{fig:kernel}
\end{figure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

% LocalWords:  LJF