### 2020-10-29 papers/types2020/pbt.tex

\section{Revisiting property-based testing for Coq}

We have presented n a previous paper~\cite{blanco19ppdp} a
proof-theoretical reconstruction of property-based testing (PBT) of
relational specifications, adopting techniques from foundational proof
certificates to account for several features of this testing
paradigm: from various \emph{generation} strategies, to
\emph{shrinking} and fault localization.

Given the connection that \textsf{coq-elpi} offers between logic
programming and the internals of Coq, it is natural to extend the
FPC-driven logic programming interpreter of the previous section to
perform PBT over \lstinline{Inductive} types.

While nothing prevents us from porting all the features we discussed
in~\cite{blanco19ppdp}, for the sake of this paper we will implement
only FPC corresponding to different flavors \emph{exhaustive}
generation and their combination, as adopted, e.g., in
SmallCheck~\cite{smallcheck} and
$\alpha$Check~\cite{cheney_momigliano_2017}.

Of course, Coq already features \textsf{QuickChick}~\cite{QChick}
(\url{https://softwarefoundations.cis.upenn.edu/qc-current}) is a
sophisticated and well-supported PBT tool, based on a different
perspective: being a clone of Haskell's QuickCheck, it emphasizes
testing executable (read \emph{decidable)} specifications with random
generators. While current research~\cite{LampropoulosPP18} aims to
increase automation, it is fair to say that testing, in particular of
relational specifications, is still very labor intensive. We do not
intend to compete with \textsf{QuickChick}, but we shall see that we
can test them immediately \lsti{Inductive} definitions that
corresponds to pure Horn programs, without any decidability proof nor
commitment to the random generation strategy, and related
configuration of generators and shrinkers.

\subsection{PBT as proof reconstruction}
\label{ssec:pbt-lp}

% \begin{metanote}
%   I'm just saying the bare minimum -am
% \end{metanote}

 If we view  a property as a logical formula
\(\forall x [(\tau(x)\wedge P(x)) \supset Q(x)]\) where $\tau$ is a
typing predicate and $P$ and $Q$ are two other predicates defined
using Horn clause specifications, providing a counter-example
consists of  negating the property, and \emph{searching} for a proof of
\[
  \exists x [(\tau(x)\wedge P(x)) \wedge \neg Q(x)]
  \tag{*}\label{eq:full}
\]
%\(\exists x [(\tau(x)\land P(x)) \land \neg Q(x)]\).

In our current setting $P$ and $Q$ will be \lsti{Inductive} propositions, while $\tau$   are honest-to-goodness datatypes. We will treat the two quite differently, in so far as elements of those types will be \emph{generated}, while predicates will only be \emph{checked}. This distinction plays also a part in interpreting the negation sign.
%
When using a focused proof system for logic extended with fixed
points~\cite{baelde12tocl}, negation corresponds to the usual intutionistic interpretation.  % proofs of formulas such as
% \[
%   \exists x [(\tau(x)\wedge P(x)) \wedge \neg Q(x)]
%   \tag{*}\label{eq:full}
% \]
% are a single \emph{bipole}: that is, when reading a proof bottom up, a
% positive phase is followed on all its premises by a single negative
% phase that completes the proof.%
% In particular, the positive phase corresponds to the \emph{generation} phase
% and the negative phase corresponds to the \emph{testing} phase.
% %

% Instead of giving a full focused proof system of a logic including
% fixed points (since, as we will argue, that proof system will not, in
% fact, be needed to account for PBT ), we offer the following analogy.
% %
% Suppose that we are given a finite search tree and we are asked to
% prove that there is a secret located in one of the nodes of that
% tree.
% %
% A proof that we have found that secret can be taken to be a
% description of the path to that node from the root of the tree: that
% path can be seen as the proof certificate for that claim.
% %
% On the other hand, a proof that no node contains the secret is a
% rather different thing: here, one expects to use a blind and
% exhaustive search (via, say, depth-first or breath-first search) and
% that the result of that search never discovers the secret.
% %
% A proof of this fact requires \emph{no} external information: instead it
% requires a lot of computation involved with exploring the tree until
% exhaustion.
% %
% This follows the familiar pattern where the positive (generate) phase
% requires external information while the negative (testing) phase
% requires none.
%
However, for the sake of PBT, we can identify a proof certificate for
$(\ref{eq:full})$ with a proof certificate for
\[
  \exists x [\tau(x)\wedge P(x)].
  \tag{**}\label{eq:short}
\]
and will resort to negation-as-failure to check that the conclusion does not hold. This also means that we will not be able to return a Coq proof terms for the refutation of our property --- and neither does \textsf{QuickChick}, which runs at the OCaml level --- although we can return the witness for the existential.
% Such a certificate would contain the witness (closed) term $t$ for the
% existential quantifier and sufficient information to confirm that
% $P(t)$ can be proved (a proof of a typing judgment such as $\tau(t)$
% is usually trivial).
% Since a proof certificate for the existence-of-a-counterexample formula
% $(\ref{eq:full})$ can be taken as a proof certificate of
% $(\ref{eq:short})$, then we only need to consider proof certificates
% for Horn clause programs.
%
%Since the logic of Horn clauses is rather simple,
% We illustrate next what such proofs and proof certificates look like
% for the rather simple logic of Horn clauses.

% \begin{metanote}
%   TODO next
%   \begin{itemize}
%   \item The elpi dep-pbt tactic
%   \item  the simple ordered example [DONE]
%   \item a bit of details about the \texttt{aexp} example
%   \end{itemize}
% \end{metanote}

\subsection{An Elpi tactic for PBT}
We will invoke the tactic in a
proof environment  where the goal is % one is trying to prove is the
the property that the system-under-test should meet. This means that, after
\lsti|intro| has been used to introduce the relevant hypotheses, we
 have a proof context % consisting of typing information for a bunch of variables
where eigenvariables  play different roles:  dependent
variables, that is the ones appearing also in the types of other variables, are subject to generation, while
non-dependent ones tags assumptions that are only to tested. This is reminiscent of bi.directional type-checking. % Obviously, the
% two are the result of and \lsti|intro| on either a dependent or a
% non-dependent product
% ; in our setting this division acquires an additional reading beyond
% what the property looks like: since specifications are encoded as pure
% Horn clauses where no dependency is involved, non-dependent variables
% simply tag the specification clauses, whereas dependent variables
% represent the typing information for the data variables that the
% specifications use.

Within the tactic the user therefore
specifies which variables of the environment should be used for
generating data and which for executing the specification. In addition
to this, the user should specify all the certificate information that
will guide the data generation phase. The goal of the current
environment will directly provide the property for which we want to
find a counterexample. % Say we want to generate a counterexample for
% some specification represented by \lsti|H1| and \lsti|H2| in the
% environment, and we want to generate the data for some variables
% \lsti|v1| and \lsti|v2| with a height bounded to 10: the final call to
% the tactic will then look like:
% \begin{lstlisting}
%   elpi pbt 10 (H1 /\ H2) (v1) (v2)
% \end{lstlisting}
Concretely, for a property
\[
\forall x_1 \colon \tau_1\dots x_n \colon \tau_m, A_1\land\dots\land A_m \supset B
\]
and the specification of a FPC, the call to the tactic will be:
\begin{lstlisting}
elpi dep_pbt <fpc> $(A_1\land\dots\land A_m) \ (x_1) \dots (x_n)$.
\end{lstlisting}
The tactic will call \lsti{check} with the given FPC on the dependent variables and will simply call a vanilla meta-interpreter (see Fig.~\ref{fig:interp} in the appendix) to test
\begin{lstlisting}
interp $(A_1\land\dots\land A_m$), not (interp $B$)}
\end{lstlisting}
where \lsti{not} in \lP's negation-as-failure operator.

To exemplify, let us add to the previous specification of list
insertion a definition of ordered list:
\begin{lstlisting}
Inductive ordered : list nat -> Prop :=
| o_n : ordered []
| o_s : forall x : nat, ordered [x]
| o_c : forall (x y : nat), forall xs,  ordered xs  -> x <= y -> ordered (x:: y :: xs).     
\end{lstlisting}
%
A property we may wish to check before embarking in a formal proof is
whether insertion preserves ordered-ness:
\begin{lstlisting}
Conjecture ins_ord:  forall (x : nat) xs rs, ordered xs -> insert x xs rs -> ordered rs.
intros x xs rs Ho Hi.
elpi dep_pbt height 5 (Ho /\ Hi ) (x) (xs).
Abort.
\end{lstlisting}
In this query the tactic tests the hypotheses \lsti{Ho} and \lsti{Hi}
against data \lsti{x,xs} generated exhaustively up to proofs bounded
by at most $5$ unfoldings of the library \lsti{Inductive} definition
of \lsti{nat} and \lsti{list}. Since we did slip in an error (where?), our
tactic reports a counter-example, namely
\verb|(* Proof Term: [0, [0; 1; 0]] *)|

In order to generate the query to the \lsti|check| predicate, some
pre-processing is needed. In particular, we need to turn the data
variables into \lP logic variables when they appear inside a
specification, and to generate queries to \lsti|check| for each of
these logic variables in association with the type of the data
variable it corresponds to.  In order to realize this pre-processing step,
we leverage extensively \lP's higher order programming features. The
substitutions are handled with the technique of \lsti|copy| clauses.

Note that the same query in \textsf{QuickChick} would have required
much more setup: if we wished to proceed relationally as above, we
would have had to provide a proof of decidability of the relevant
notions. Were we to use functions, then we would have to implement a
generator and shrinker for ordered lists.

\todo{the \texttt{aexp} example}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

% LocalWords:  Bedwyr bipole elpi QuickCheck OCaml TODO