### 2020-11-01 papers/types2020/conc.tex
related work bits spear thru the paper
ported the appendix and sec3 to kernel-v2

sec3 is still a mess

\section{Conclusion and future work}

\todo{sum up}

\begin{metanote}
  Decided against a related work section: most is already discussed in the intro, the rest can be mentioned in passing --am
\end{metanote}
% \subsection{Related work}
% \label{ssec:rel}

% \begin{itemize}
% \item Brief discussion/comparison with \texttt{auto}
% \item Mention other papers by Enrico as applications of coq-elpi~\cite{coen19mscs,tassi19itp}\dots [done]
% \item Say we \textbf{do} not talk about Coq metaprogramming (Ltac, Mtac), do say that doing our to application in MetaCoq~\cite{sozeau2020metacoq} would be a disaster (reimplement unfication, backtracking, binders) [done]
% \end{itemize}

\subsection{Future work}
\label{ssec:fut}

The code reported in Fig.~\ref{fig:augmented} is somewhat idealized in
so far that we omitted a rule performing weak-head reduction (allowing
to handle directly existential goals) and more importantly that used
the \emph{delay} mechanism. This a distinguished feature of Elpi, whose
usefulness has been advocated in ~\cite{DunchevCT16,coen19mscs}.  The
use of constraints for data generation is well
developed~\cite{FioravantiPS15} and we could try to leverage to
improve PBT using partially instantiated terms: for example generating
say lists with at least two elements as \texttt{[1,2|Tail]}, without
recurring to needed narrowing as in in
LazySmallCheck~\cite{smallcheck}.

\todo{Link to the repo? \url{https://github.com/proofcert/fpc-elpi}}

Extending our tactic to handle computations (\texttt{simp}) and/or to
return the partial proof state, unsolved goals in particular. Mention
\emph{Harpoon}, Beluga's new tactic language?

Extending the LJF kernel to handle defined atoms
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

% LocalWords:  Elpi