### 2020-11-03 papers/types2020/intro.tex

\section{Introduction}
\label{sec:intro}

Recently, C. Sacerdoti Coen and E. Tassi developed the Elpi
implementation \cite{dunchev15lpar} of \lP \cite{miller12proghol} and
more recently, Tassi has made Elpi available as the \texttt{coq-elpi} plugin to the Coq
proof assistant \cite{tassi18coqpl} (\url{https://github.com/LPCIC/coq-elpi}).  This implementation of \lP
extends earlier ones in primarily two directions:  First,
Elpi adds a notion of constraints and constraints handling rules, which makes it more expressive than
the Teyjus implementation \cite{nadathur99cade} of \lP. Second, the
plugin version of Elpi comes equipped with  a quotation and anti-quotation syntax for mixing
Coq expressions with \lP program elements and an API for accessing the
Coq environment, typechecking included.

The logic programming interpreter that underlies Elpi provides several
convenient features for the kind of meta-programming tasks that can
arise within modern proof assistants.  For example, \lP provides a
declarative and direct treatment of abstract syntax that contains
bindings, including capture-avoiding substitution,
unification, and recursive programming.  Elpi spares the
programmer from dealing with low-level aspects of bindings
representation (e.g., De Bruijn indexes) while still having clean and
effective ways to manipulate binding structures.

Since relations (not functions) are central in \lP, Elpi is
capable to providing direct support for the many relations that have a
role in proof assistants implementation and usage.  Such relations include typing (e.g.,
$\Gamma\vdash M\colon\sigma$), evaluation (e.g., natural semantics
specifications \cite{kahn87stacs,hannan93jfp}), and interaction (e.g.,
structured operational semantics \cite{plotkin81,milner89book}).

Felty has also
made the point that LCF-style tactics and tacticals can be given an
elegant and natural specification using the higher-order relational
specifications provided by \lP \cite{felty93jar}.  Some recent
implementations built using Elpi
support the usefulness of
higher-order logic programming as a meta-programming language 
for proof assistants in general~\cite{DunchevCT16,coen19mscs} and in particular
for Coq via the \texttt{coq-elpi} plug in~\cite{tassi19itp,CohenST20} .
%

In this paper, we present two applications of Elpi within Coq. With
these examples, we shall illustrate that Elpi is useful not only
because of its meta-programming features,
% (e.g., a builtin treatment of bindings in syntax)
but also because it soundly
implements a higher-order intuitionistic logic.
\todo{Explain better? Our kernels are deep embedding, aren't they? -am}The latter aspect of
Elpi makes it possible to encode the proofs and proof theory of
specific subsets of the Coq logic.

While other meta-programming frameworks based on functional
programming such as \emph{MetaCoq}~\cite{sozeau2020metacoq} can and have been used for related endeavors, we believe (together with~\cite{DunchevCT16}) that they would require much more boilerplate.

Before we can present these examples, we first highlight the rather
striking differences in notions of computing and reasoning that arise
on each side of the \texttt{coq-elpi} API.   We will also present a quick
summary of the key proof theory concepts that are used by our example
applications.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

% LocalWords:  Sacerdoti Coen Tassi Elpi Teyjus Felty LCF tacticals
% LocalWords:  intuitionistic elpi coq MetaCoq