### 2020-11-05 papers/types2020/intro.tex

\section{Introduction}
\label{sec:intro}

Recently, C. Sacerdoti Coen and E. Tassi developed the Elpi
implementation \cite{dunchev15lpar} of \lP \cite{miller12proghol} and
more recently, Tassi has made Elpi available as the Coq-Elpi
plugin~\cite{tassi18coqpl} (\url{https://github.com/LPCIC/coq-elpi})
to the Coq proof assistant.  This implementation of \lP extends
earlier ones in primarily two directions: First, Elpi adds a notion of
constraints and constraints handling rules, which makes it more
expressive than the Teyjus implementation \cite{nadathur99cade} of
\lP. Second, the plugin version of Elpi comes equipped with a
quotation and anti-quotation syntax for mixing Coq expressions with
\lP program elements and an API for accessing the Coq environment,
type checking included.

The logic programming interpreter that underlies Elpi provides several
convenient features for the kind of meta-programming tasks that can
arise within modern proof assistants.  For example, \lP provides a
declarative and direct treatment of abstract syntax that contains
bindings, including capture-avoiding substitution,
unification, and recursive programming.  Elpi spares the
programmer from dealing with low-level aspects of bindings
representation (e.g., De Bruijn indexes) while still having clean and
effective ways to manipulate binding structures.

Since relations (not functions) are central in \lP, Elpi is capable to
providing direct support for the many relations that have a role in
proof assistants implementation and usage.  Such relations include
typing (e.g., $\Gamma\vdash M\colon\sigma$), evaluation (e.g., natural
semantics specifications \cite{kahn87stacs,hannan93jfp}), and
interaction (e.g., structured operational semantics
\cite{milner89book,plotkin81}).

Felty has also made the point that LCF-style tactics and tacticals can
be given an elegant and natural specification using the higher-order
relational specifications provided by \lP \cite{felty93jar}.  Some
recent implementations built using Elpi support the usefulness of
higher-order logic programming as a meta-programming language for
proof assistants in general~\cite{DunchevCT16,coen19mscs} and, in
particular, for Coq via the Coq-Elpi plug~\cite{CohenST20,tassi19itp} .

In this paper, we present two applications of Elpi within Coq. With
these examples, we shall illustrate that Elpi is useful not only
because of its meta-programming features, but also because it soundly
implements a higher-order intuitionistic logic: such implementations
of higher-order logic have long been known to provide powerful and
flexible approaches to implementing many different logics and their
proof systems \cite{felty89phd,felty88cade,paulson89jar}.  Following
that tradition, the Elpi system makes it possible to encode the proofs
and proof theory of various subsets of the logic behind Coq (see also
\cite{felty93lics}).

While other meta-programming frameworks based on functional
programming such as \emph{MetaCoq}~\cite{sozeau2020metacoq} can and
have been used for related endeavors, we believe (together
with~\cite{DunchevCT16}) that they would require much more
boilerplate code.

Before we can present these examples, we first highlight the rather
striking differences in notions of computing and reasoning that arise
on each side of the \texttt{coq-elpi} API.   We will also present a quick
summary of the key proof theory concepts that are used by our example
applications.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

% LocalWords:  Sacerdoti Coen Tassi Elpi Teyjus Felty LCF tacticals
% LocalWords:  intuitionistic elpi coq MetaCoq