### 2021-03-05 papers/types2020/revisions/main.tex
Improved the abstract.

%% Important dates:
%%  Paper submission: 19 October 31 October 2020
%%  Author notification: 18 January 2021
%%  Final version:  15 February 2021
%%  Publication (presumably): end of March 2021

%% Style requirements of LIPIcs.  The recommended length of a paper is
%% 12-15 pages, excluding front-page(s) (authors, affiliation, keywords,
%% abstract, ...), bibliography and an appendix of max 5 pages. If you
%% need more pages, please ask the editors.

%% Papers have to be submitted in pdf through EasyChair:
%% https://easychair.org/conferences/?conf=types2020postproceed

\documentclass[a4paper,USenglish,cleveref, autoref, thm-restate]{lipics-v2019}

\input{../head}
\input preamble

\title{Two applications of logic programming to Coq} 
\titlerunning{Two applications of logic programming to Coq}

\begin{document}
\maketitle

\begin{abstract}
The logic programming paradigm provides a flexible setting for
representing, manipulating, checking, and elaborating proof
structures.  This is particularly true when the logic programming
language allows for bindings in terms and proofs.  In this paper,
we make use of two recent innovations at the intersection of logic
programming and proof checking.  One of these is the
\emph{foundational proof certificate} (FPC) framework which provides a
flexible means of defining the semantics of a range of proof
structures for classical and intuitionistic logic.  A second
innovation is the recently released Coq-Elpi plugin for Coq in which
the Elpi implementation of \lP can send and retrieve information to
and from the Coq kernel.  We illustrate the use of both this Coq
plugin and FPCs with two example applications.  First, we implement an
FPC-driven sequent calculus for a fragment of the Calculus of
Inductive Constructions and we package it into a tactic to perform
property-based testing of inductive types corresponding to pure
(higher-order) Horn programs.  Second, we implement in Elpi a proof
checker for first-order intuitionistic logic and demonstrate how proof
certificates can be supplied by external (to Coq) provers and then
elaborated into the fully detailed proof terms that can be checked by
the Coq kernel.
\end{abstract}

\input intro
\input cultures
\input sec3
\input pbt
\input elab

\input conc
\bibliography{../l}
%\newpage
\appendix
\input app
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

% LocalWords:  pbt elab conc Elpi LJF FPC Coq sequent provers