### 2021-03-05 papers/types2020/revisions/cover.tex

\documentclass{article}

\begin{document}
\title{Cover letter for ``Two applications\dots''}
\maketitle{}

\section*{Main Changes}

\begin{itemize}
\item ``auto'' with conversion
\item ``auto'' with Hints (?)
\item other \dots
\end{itemize}
\section*{Reviewer 1}
\begin{verbatim}
An introduction to the topic, or at
least a section explaining the application in the paper, would be
necessary to open the paper to a larger audience.
\end{verbatim}
comment
\begin{verbatim}
It is misterious why you choosed to only use constructors: you never
justify this restriction. There is also an hint that you can use the Hints
that eauto uses, but this is not reflected in your code. How can it be?
\end{verbatim}
comment
\section*{Reviewer 2}

\end{document}